---
title: "Data Manipulation"
author: "Ashley Kang"
date: "9/18/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## Import FAS csv files

Import my first csv

```{r}
litters_data = read_csv(file = "./data/FAS_litters.csv")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv(file = "./data/FAS_pups.csv")
# pups_data = read_csv(file = "~/Desktop/data_wrangling/data/FAS_pups.csv")
pups_data = janitor::clean_names(pups_data)

#first step = "print"" out data-set 
# in console, type dataset name and check"
```

```{r}
select(litters_data, group, litter_number, gd0_weight)


select(litters_data, litter_number, gd0_weight, pups_survive)
#tell select exactly what variables you're interested in
```

## specify range of variables

```{r}
select(litters_data, gd_of_birth:pups_survive)
## selecting evertyhing from gd of birth to pups survive

select(litters_data, group, gd_of_birth:pups_survive)
```

## can specify what you don't want to keep
```{r}
select(litters_data, -litter_number)
# keeping everything except litter number
```

## rename gd0_weight to gest_day_0_weight
```{r}
select(litters_data, group, litter_number, gest_day_0_weight = gd0_weight)
# select is designed to make it easy for you to select variables you care about
rename(litters_data, gest_day_0_weight = gd0_weight)
# rename is helpful if you have 1 or 2 variables you want to rename
```

## using select_helpers
```{r}
select(litters_data, starts_with("gd"))
# extracts only gd variables

select(litters_data, litter_number, starts_with("pup"))
# mix and match criteria

select(litters_data, litter_number, everything())

```

## pups data
```{r}
select(pups_data, litter_number, sex, pd_ears)
```

## filter observations
```{r}
filter(litters_data, gd0_weight < 25)
filter(litters_data, gd0_weight >= 25)

filter(litters_data, pups_born_alive == 8)
filter(litters_data, is.na(gd0_weight))
# is.na = is there missing data in column?

filter(litters_data, !is.na(gd0_weight))
# !is.na = opposite of is.na

filter(litters_data, group == "Low7")

filter(litters_data, group == "Low7" | group == "Low8")
# | means or; jeff doesn't like or it gets weird

filter(litters_data, group %in% c("Low7", "Low8"))

filter(litters_data, pups_born_alive %in% 4:6)

filter(litters_data, pups_born_alive %in% 4:7, !is.na(gd0_weight))
#give me all of the data where first and second condition is true
```

## filter example

```{r}
filter(pups_data, sex == 1)
filter(pups_data, pd_walk < 11, sex == 2)
```

## mutate

create variable

```{r}
mutate(litters_data,
        wt_gain = gd18_weight - gd0_weight,
       group = tolower(group)
)
#can do multiple variable mutations within one mutate
#first, creating new variable wt_gain based on function of existing variables
#next, overwriting existing group variable to a new version (makes everything lowercase)

mutate(litters_data,
      wt_gain = gd18_weight - gd0_weight,
      wt_gain_squared = wt_gain^2
)

```

## Arrange

Arrange the data:
```{r}
arrange(litters_data, gd0_weight)
# gd0_weight is arranged smallest 

arrange(litters_data, pups_born_alive, gd0_weight)
# pups_born_alive is arranged first (smallest to largest) and then gd0_weight

arrange(litters_data, desc(pups_born_alive), gd0_weight)
# descending order
```

## Piping!!!!

Look at intermediate object approach:
```{r}
litters_data_raw = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data_clean_names = janitor::clean_names(litters_data_raw)
litters_data_selected_cols = select(litters_data_clean_names, -pups_survive)
# removing pups_survive
litters_data_with_vars = mutate(litters_data_selected_cols, 
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group))
litters_data_with_vars
```

Use nested function calls:

```{r}
litters_data_clean = 
  mutate(
    select(
      janitor::clean_names(
        read_csv("./data/FAS_litters.csv", col_types = "ccddiiii")
        ), 
    -pups_survive
    ),
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group)
  )
litters_data_clean
```

Use actual piping:

```{r}
read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>% 
  janitor::clean_names() %>% 
  select(-pups_survive) %>% 
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = tolower(group)
  ) %>% 
  filter(.data - ., !is.na(gd0_weight))

# 1. read in data set
# hot key: shift+command+m=hot key for pipe
# %>% passing in first argument to next
```


Illustrate placeholder use with "lm" (linear model)

```{r}
read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>% 
  janitor::clean_names() %>% 
  select(., -pups_survive) %>% 
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = tolower(group)
  ) %>% 
  filter(!is.na(gd0_weight)) %>% 
  lm(gd18_weight ~ gd0_weight, data = .)
# whatever dataframe resulted prior to the pipe operator in filter will exist right there
# for lm we have to put data = . because the first argument isn't data like it is for
# select or mutate which already assumes that the first argument is data, so we don't have 
# to be explicit in select or mutate but we do for lm
```

```{r}
clean_data = read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>% 
  janitor::clean_names() 

select(clean_data, -pups_survive) %>% 
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = tolower(group)
  ) %>% 
  filter(!is.na(gd0_weight)) %>% 
  lm(gd18_weight ~ gd0_weight, data = .)
# what pipe does is instead of naming clean_data and plugging it in into the select
# it simplifies that step and just assumes
```

